---
title: "SML Assignment - 03"
author: "MADHUSUDHAN ASHWATH - 19203116"
date: "4/19/2020"
output: word_document
---

# Library files

```{r, warning= FALSE }

# Install the packages
install.packages("mlbench")
install.packages("nnet")
install.packages("randomForest")

```

```{r, warning= FALSE }

# Load the packages
library(mlbench)
library(nnet)
library(randomForest)
```

# Setting up data

About Data Set :
1.	The data set consists of 37 variables and 6435 observation.
2.	Multi-spectral values of pixels n 3x3 neighborhood’s in a satellite image.
3.	The images are classified into 6 types
  a.	cotton_crop
  b.	damp_grey_soil
  c.	grey_soil
  d.	red_soil
  e.	soil_with_vegatation_stubble
  f.	very_damp_grey_soil

```{r}
# Import data
data("Satellite")

# Removing spaces and factorizing
Satellite$classes = gsub(" ", "_", Satellite$classes)
Satellite$classes = factor(as.character(Satellite$classes))
```

# Splitting of data, Seting the trail

```{r}

# Setting seed
set.seed(777222)

# Number of rows
D = nrow(Satellite)

# Sampling the data
keep = sample(1:D, 5500)
test = setdiff(1:D, keep)
dat = Satellite[keep,]
dat_test = Satellite[test,]

# Converting to factors
dat$classes = factor(dat$classes)
dat_test$classes = factor(dat_test$classes)

# Row counts

d_count = nrow(dat)
dt_count = nrow(dat_test)

# Trails

trail = 100

# Matirx to store the output

mat = matrix(NA , trail , 4)
mat = as.data.frame(mat)

# Changing the column name for ploting
colnames(mat) = c("Multinomial" , "RandomForest", "BestFit" , "Test")


```

Procedure: 

1.	Split the data into 3 parts.
  a.	Training set
  b.	Test Set
  c.	Validation Set
2.	Test which model is more accurate Multinomial Regression or Random forest.
3.	Both the models are fitted to training data to predict the model
4.	Validation data set is used to determine the accuracy between the models.
5.	The model with the highest accuracy will be the best.
6.	The above procedure is ran for 100 trails.
7.	The accuracy is noted in a form of matrix(in our analysis).
8.	The summary of the best model with good accuracy is obtained.
9.	Box plot is shown to see the data summary.
10.	Later plot a graph of accuracy v/s the trails for both the model.

# Model Building, Classifier Comparission

```{r}
# Looping for all the trails
for (n in 1 : trail) 
  {
  # training and validation sets
  print(paste0("Trials :", n))
  
  ## 75% of data as training set
  dat_train = sample(1:d_count, size = 0.75*d_count)  
  
  ## 25% of data as Validation set
  dat_val = sample(setdiff(1:d_count, dat_train), size = 0.25*d_count)
  
  # Model Building
  
  ## Multinomial Regression
  fit = multinom(classes ~ ., data = dat, subset = dat_train, maxit = 500)
  
  ## Random Forest
  fitrf = randomForest(classes ~ ., data = dat, subset = dat_train, importance = TRUE)
 
  # Ovservation of Validation data
  
  ## Multinomial Regression
  pred = predict(fit, newdata =  dat[dat_val,])
  tab = table(dat$classes[dat_val], pred)
  accmr = sum(diag(tab))/sum(tab)

  # Random Forest
  predrf = predict(fitrf,newdata =  dat[dat_val,])
  tabrf = table(dat$classes[dat_val], predrf)
  accrf = sum(diag(tabrf))/sum(tabrf)
 
  # Storing onto matrix
  acc = c(Multinomial = accmr , RandomForest = accrf)
  mat[n,1] = accmr
  mat[n,2] = accrf
  
  
  # Calculationg of accuracy
  BestFit = names(which.max(acc))
  switch(BestFit,
         ## Multinomial Regression
         Multinomial = 
           {
           predtest = predict(fit, newdata = dat_test)
           tabtest = table(dat_test$classes, predtest)
           accbest = sum(diag(tabtest))/sum(tabtest)
           },
         ## Random Forest
         RandomForest = 
           {
           predtestrf = predict(fitrf, newdata = dat_test)
           tabtestrf = table(dat_test$classes, predtestrf)
           accbest = sum(diag(tabtestrf))/sum(tabtestrf)
          }
  )
  mat[n,3] = BestFit
  mat[n,4] = accbest
}

# Tabulating the result

table(mat[,3])
tapply(mat[,4], mat[,3], summary)

```

# Box Plot

```{r}
# Plotting Box plot to observe the distrubution
# Box plot
boxplot(mat$Test ~ mat$BestFit)
stripchart(mat$Test ~ mat$BestFit , add = TRUE, vertical = TRUE, method = "jitter", pch = 15, col = adjustcolor("red", 0.5))

```

# standard deviation for plot

```{r}
average = t(lapply(mat[,1:2], mean))
print("Average")
average
std = apply(mat[,1:2], 2, sd)/sqrt(trail)
std = as.matrix(std)
print("Standard Deviation")
std

```

# Plot of Multinomial Regression v/s Random  Forest

```{r}
# plot the values
matplot(mat[,1:2], type = "l", lty = c(2,3), col = c("red","purple"), xlab = "Trails-Samples", ylab = "Accuracy",main = "Multinomial Regression v/s Random  Forest", ylim = c(0.80, 0.95), xlim = c(0, 100))

# Mean Line
abline(h = average, col = c("red", "purple"))

# Setting the intervals
upper = rep(c(average[[2]]-2*std[2], average[[2]]+2*std[2]), each = trail)
lower = rep(c(average[[1]]-2*std[1], average[[1]]+2*std[1]), each = trail)
polygon(c(1:trail, trail:1), lower, col = adjustcolor("red", 0.2), border = FALSE)
polygon(c(1:trail, trail:1), upper, col = adjustcolor("purple", 0.2), border = FALSE)

# Legend
legend("bottomright", fill = c("purple" , "red"),
       legend = c("Random forest" , "Multinomial regression"), bty = "y")

```

Analysis:

1.	It is seen that the accuracy of the random forest model is the best as it has an accuracy of 0.9523.
2.	The accuracy of the model varied from 0.932 to 0.9701.
3.	Thus can conclude that the random forest model has a high classification rate. 
4.	Seeing the performance of the classifier, can see that the mean predictive performance of random forest is above 90%, which implies high predictive performance.
5.	The prediction speed is significantly faster than training speed because the generated values can be saved for future uses.
6.	It is capable of handling outliers by binning them.
7.	Minimizes the overall error rate for unbalanced dataset.
8.	Based on all the above points we can conclude random forest is the best model with good accuracy and better predictor overall.

